The Heavy Photon Search Test Detector

### 2013-11-11 hps-testrun-nim.tex
Stepans changes to mostly ecal perf.

\section{Electromagnetic Calorimeter}
\label{sec:ecal}

@@

\subsection{Purpose and Design Requirements}

The electromagnetic calorimeter (ECal) provide the trigger for the data acquisition and will be used for 
electron identification during the data analysis. The ECal is positioned after the analyzing dipole 
magnet and cover the full acceptance region of the SVT. It consist of two halves, split in the horizontal 
beam plane. The gap between upper and lower parts, $XXXX$~mrad (15mrad for electron running),  
as seen from the target location, is necessary to avoid the "dead zone" where occupancy and radiation 
prevents any instrumentation to be situated.

The energy of electrons of interest will be in the range 
$0.5-6.5$~GeV. The ECal modules must have sufficient radiation lengths to absorb the full energy of these 
scattered electrons, have fine enough granularity to handle a high rate of electromagnetic 
background and survive large doses of radiation, especially near the beam plane. In addition, it needs to 
have excellent hit time resolution, $<10$~ns, in order to beat down beam backgrounds.

The compact design requirements: high radiation environment , high rate and the presence of 
high magnetic field, are fulfilled by a lead-tungstate (PbWO$_{4}$) crystal calorimeter with magnetic field 
resistant avalanche photodiode (APD) readout.  Such crystals has been operational in the inner 
calorimeter of the CLAS experiment~[XXXX] at JLab since 2005 and meet all requirements set by 
HPS. These crystals were subsequently repurposed for the ECal.

@@

%Step\subsubsection{PbWO$_4$ Crystals}
The HPS test run ECal uses PbWO$_{4}$ crystals, APDs, and preamplifier boards from the inner calorimeter (IC) 
of the CLAS detector~\cite{2008erhm.conf..421N}.  In Fig.~\ref{fig:ecal-module} a schematic view of the module 
assembly is shown. Lead-tungstate crystals are $160$ mm long and are tapered with front face of 
$13.3\times13.3$~mm$^2$, tapered to $16\times16$~mm$^2$ at the rear face. The crystals were fabricated with 
very tight tolerances: $19~\mu$m for 13.3~mm dimension and $13~\mu$m for the 16~mm dimension. Each crystal 
is wrapped in VM2000 multilayer polymer mirror film.
The APDs (Hamamatsu S8664-55) with $5\times5$~mm$^2$ active area and $75\%$ quantum efficiency were 
glued on the rear face of the crystals using MeltMount 1.7 thermal plastic adhesive. The low gain of APDs ($\sim 
200$) was compensated with custom made preamplifier boards, which provide factor 2000 amplification of the 
ADP signal. 
\begin{figure}[]
\begin{center}
{\small
\includegraphics[width=7cm]{figures/ecal-module-schematic.png}
\caption{Schematic vied of an ECal module.}
\label{fig:ecal-module}
}
\end{center}
\end{figure}

@@

During the test run ECal was mounted downstream of the analyzing dipole magnet at the distance of about 
$147$~cm from the 
upstream edge of the magnet. 
%The two ECal modules are positioned just above and below the ECal vacuum chamber as shown in bottom panel of Fig.~\ref{fig:ecal}. 
%The primary electron beam, scattered electrons and bremsstrahlung photons pass through the system in vacuum. 
The gap between innermost edge of the crystals of two parts was just $7.4$~cm from the beam. The temperature 
stability of the system during the test run was better than $1^\circ~$F, the temperature uniformity inside the box 
was better than $3^\circ~$C.

@@

\subsection{Signal readout}
\label{sec:fadc}
Amplified APD signal from the motherboard is sent to a $2:1$ signal splitter. After the split, 2/3 of the signal is sent 
a discriminator then to a Time-to-Digital Converter. The 1/3 of the signal is fed to a single channel of 16-channel 
JLab Flash Amplitude-to-Digital Converter VXS module (FADC250)~\cite{fadc} shown in Fig.\ref{fig:fadc}.  
\begin{figure}[t]
\begin{center}
{\small
\includegraphics[width=7cm]{figures/FADC250_Photo_001.jpg}
\caption{A JLab FADC250 VXS module.}
\label{fig:fadc}
}
\end{center}
\end{figure}
Two 20-slot VXS crates with 14 FADC bards were needed to accommodate the system: one for each half of the 
ECal with 221 channels.

The FADCs store 12-bit digitized samples at 250~MSPS in 8~$\mu$s deep pipelines. 
When a trigger is received, the appropriate part of the pipeline is accessed for the readout. HPS used FADCs in 
the co-called integration mode. A predefined number of samples of the signal that exceeds a predefined threshold 
have been summed and stored in 17-bit register for readout. The number of samples for integration was defined 
by two parameters that could be set separately, the number of samples integrated before the signal crossed 
threshold ($NSB$) and  the number of samples integrated after the signal crossed threshold ($NSA$), 
see Fig.~\ref{fig:hps_trigger_data}. 
This scheme significantly compresses the data input 
to the FADC. During data analysis, a pedestal value is subtracted to obtain the actual summed energy.
\begin{figure}[]
\begin{center}
{\small
\includegraphics[width=7cm]{figures/hps_trigger_data}
\caption{\small{FADC data paths}}
\label{fig:hps_trigger_data}
}
\end{center}
\end{figure}

@@

The HPS test run trigger system is designed to efficiently select \ee{} events by 
using information from the ECal . The trigger looks for time coincidences of clusters in the top and bottom half 
of the ECal.  The trigger system can be broken down into three sections (see Fig.~\ref{fig:hps_trigger_cal}):
 \begin{itemize}
 \item FADC (pulse finding): samples and digitizes the signal pulses from each detector channel. Sends the 
 measured pulse energy and arrival time to the Crate Trigger Processor (CTP),
\item CTP (cluster finding): groups FADC pulses from each half of the ECal into clusters. The cluster energy and 
arrival time are sent to the Sub-System Processor (SSP),
 \item SSP (cluster pair finding): searches for time coincidence of pairs of clusters from the top and bottom half of 
 the ECal and applies topological selections.
 \end{itemize}
 \begin{figure}[t]
\begin{center}
{\small
 \includegraphics[width=7cm]{figures/hps_trigger_cal}
\caption{Block diagram of the ECAL trigger system consisting of the FADC that samples and digitizes the detector 
	channel signals and sends them for cluster finding in the CTP. The CTP clusters are sent to the SSP where 
	the final trigger decision is taken based on pairs of clusters in both halves of the ECal. The decision is sent to
	the Trigger Supervisor (TS) that generates the necessary signals to readout the sub-detectors.}
 \label{fig:hps_trigger_cal}
}
\end{center}
 \end{figure}
The time coincidence window of pairs of clusters in the top and bottom half of the ECal are programmable 
with 4~ns resolution.

As described above, the first stage of the trigger logic is incorporated into the 
FPGA's on the FADC boards. For the trigger path, every channel has, in addition to $NSB$ and $NSA$: 
\begin{itemize}
% \item number of samples integrated before the threshold crossing (NSB),
 %\item number of samples integrated after the  threshold crossing (NSA),
 \item trigger threshold, measured in ADC counts, 
 \item a pedestal, 
 \item a conversion factor (gain) that converts the ADC counts to energy in MeV (with 13 bits:  from 0 to 8191~MeV), 
 \item an energy discriminator threshold (minimum energy cutoff).
 \end{itemize}
Note that the threshold for the trigger path can be set independently from the readout threshold.
The pedestal value is subtracted from the integrated sum over $NSB+NSA$ samples and 
converted to MeV units using a supplied gain conversion factor. The energy discriminator can be used to cut 
off low energy pulses before reporting to CTP. 
The values reported to the CTP are the 5-bit pulse energy and 3-bit time information at which the 
pulse crossed the threshold. Data for every channel is sent to CTP every 32~ns. With the available 
3-bit time information, the CTP looks for time coincidence of crystal signals with 8~ns resolution (limit is 4~ns).
The cluster finding algorithm is fast and makes use of the parallel processing nature of FPGA's by 
simultaneously searching for 125 clusters, up to 3x3 in size, across the calorimeter crystal array and performs 
the following tasks:
\begin{itemize}
\item Adds energy from hits together for every 3x3 square of channels in ECal.
\item Hits are added together if they occur (leading edge) within a programmable number of 4~ns clock cycles of each other (HPS test run used 8~ns coincidence time interval).
\item If the 3x3 energy sum is larger than the programmable cluster energy threshold and the sum is greater than any neighboring 3x3 windows, the CTP reports the cluster parameters to the SSP. 
\end{itemize}
The CTP evaluates all hits in its half of the calorimeter every 4~ns. A programmable time window is used to 
allow hits that are out of time with each other to be considered as part of a cluster sum. This is done by 
reporting hits when they occur and then reporting them again for the next $N$ number of 4~ns clock cycles, 
where $N \in [0,7]$. This is useful to deal with skew and jitter that develop from the detector, cabling, and 
electronics. As described above, the CTP only selects the 3x3 window with the highest energy sum of its 
neighbors. This filtering is applied to deal with overlapping clusters and cases where the cluster is larger than 
a 3x3 window.

@@

During the test run, for the most part trigger system run in single cluster mode and was tested to work for up to 
20~kHz trigger rates.

\subsection{Data Acquisition and Online Computing}
\label{sec:daq}
For the ECal, every VXS crate contains a Readout Controller (ROC)
that collects digitized information, processes it, and sends it on to the Event Builder (EB). The ROC is a single 
blade Intel-based CPU module running DAQ software under CentOS Linux OS. For the SVT DAQ system, 
the ROC application runs on a dedicated DAQ desktop computer. The EB assembles 
information from the SVT and Cal ROCs into a single event which is passed to the Event 
Recorder (ER) that writes it to a RAID5-based data storage system capable of handling up to 100~MB/s. The 
EB and other critical components run on multicore Intel-based multi-CPU servers. The DAQ network system is 
a Foundry router providing high-speed connections between the DAQ components and to the JLab 
computing facility. The SVT ROC, which must handle large data volumes, has a 10~Gb/s link to the Foundry 
router, while a 1~Gb/s link is adequate for the ECal. A 10~Gb/s uplink to the JLab 
computing facility is used for long-term storage.

@@

%\subsubsection{ECal Calibration and Performance}

Only 385 out of 442 modules (87\%) were operational during the test run.  Of 442 modules, 39 were disabled or 
disconnected and were not read out by the DAQ. 
13 of these were not read out because of a shortage of FADC readout boards.
The remainder either had no HV bias on the APD, or were disabled in the FADC software due to noise.
In the data, we identified two types of abnormal channels. 
One FADC was not sending trigger signals correctly, resulting in low efficiency.
 This affected the 13 channels read out by that FADC. 5 channels were diagnosed as noisy because they 
 had a high incidence of hits out of coincidence with the trigger. A large number of channels were originally 
 misidentified as noisy because they had much higher hit occupancy than neighboring channels.
Gain calibration shows that these channels have high gain (and thus lower energy threshold) but are otherwise 
normal. The abnormal channels were ignored in analysis in order to simplify comparison with Monte Carlo. This 
leaves 385 useful channels---87\% of the ECal.

The noise level of the ECal modules were measured to be approximately 15~MeV. The ECal was operated 
with an effective readout threshold of 70~MeV. Modules above thresholds are clustered using a simple robust 
algorithm described in Sec.~\ref{sec:trigger}.

%\subsubsection{Calibration}
%\label{sec:ecal_calibration}
The noise and pedestal of the readout chain are calibrated by running the ECal readout in a mode where the 
preamplifier output is sampled every 4~ns in a time window of 100 samples: by looking at a part of the 
window before the hit,  the readout channel pedestal can be obtained. 
We calibrate gain of the individual ECal channels using the SVT measurement of track momentum and 
comparison to Monte Carlo simulation.  We disable all SVT and ECal channels in the simulation that were 
inoperable or noisy in the test run, so any efficiency or bias effects that affect the real data should be reflected 
in the simulation as well; then we use a formula to compute the ``weighted E/p'' for a crystal, representing the average E/p for clusters that include the crystal
% $\frac{\sum_j w_{j,i}}{\sum_j\frac{P_j}{E_j}w_{j,i}}$
, and iteratively adjust the gains until the weighted E/p is equal for test run data and simulation.
\begin{figure}[]
\begin{center}
{\small
	\includegraphics[width=7cm]{figures/eoverp_simulation}
	\caption{Weighted E/p from Monte Carlo simulation after gain calibration.}
	\label{fig:gains}
}
\end{center}
\end{figure}
These gains can then be used to convert from ADC counts in a channel to the energy deposited into that ECal 
crystal.
The other information needed to find the energy of an incident particle is the sampling fraction---the ratio of 
energy read out from crystals to energy of an incident particle. The conventional sampling fraction, 
the fraction of incident energy that is deposited in crystals, is approximately 0.9  for the center of the ECal and less 
at the 
edges. For our readout, there is additional energy lost because crystals under the readout threshold are not 
read out. The weighted E/p used in calibration (see Figure \ref{fig:gains}) is an approximate measurement of sampling fraction, but the sampling fraction is energy-dependent because of the effect of readout threshold. 
A full computation of sampling fraction can be done using simulation.

@@

To eliminate trigger bias in checking the trigger decision, we use a tag and probe method. To check trigger 
performance in one half of the ECal, we tag events where there was a trigger in the other half, and exactly one 
probe cluster in the ECal half under test. 
We then measure trigger efficiency (proportion of tagged events where there was a trigger) as a function of 
ADC counts and energy of the probe cluster.
These turn-on curves are shown for the top half of the ECal in Fig.~\ref{fig:turnon}. 
\begin{figure}[ht]
\begin{center}
{\small
	\includegraphics[width=7cm]{figures/top_turn_on__bottom_tag}
	\includegraphics[width=7cm]{figures/top_turn_on__bottom_tag_E}
	\caption{Trigger turn-on as a function of probe cluster ADC counts (top) and probe cluster energy in MeV
	(bottom). Both plots are for the top half of the ECal; bottom is similar. Energy is not corrected for sampling 
	fraction.}
\label{fig:turnon}
}
\end{center}
\end{figure}
The trigger threshold is seen to be 1280 ADC counts as expected. The threshold is not perfectly sharp in this 
analysis because of uncertainties in the conversion from readout to trigger hits but based on comparisons 
with Monte Carlo simulation we believe the trigger worked exactly as specified. 
The trigger threshold in terms of cluster energy is very uneven for two reasons; gain variations between 
different ECal crystals lead to threshold variations and the nonlinearity of the time-over-threshold integral 
means that the effective threshold is higher for clusters that span multiple crystals.
Overall the trigger appears to have functioned exactly as intended. Changes planned for the next run 
(constant integration window and per-crystal gain calibration constants for the trigger) will solve both of the 
issues that led to threshold variations in the test run.

>>>>

\section{Electromagnetic Calorimeter}
\label{sec:ecal}
\section{Electromagnetic Calorimeter}
\label{ecal}

@@

@@

%Step\subsubsection{PbWO$_4$ Crystals}
The HPS test run ECal uses PbWO$_{4}$ crystals, APDs, and preamplifier boards from the inner calorimeter (IC) of the CLAS detector\cite{2008erhm.conf..421N}.  In Figure~\ref{fig:ecal-module} a schematic view of the module assembly is shown. Lead-tungstate crystals are $160$ mm long and are tapered with front face of $13.3\times13.3$~mm$^2$, tapered to $16\times16$~mm$^2$ at the rear face. The crystals were fabricated with very tight tolerances: $19~\mu$m for 13.3~mm dimension and $13~\mu$m for the 16~mm dimension. Each crystal is wrapped in VM2000 multilayer polymer mirror film.
The APDs (Hamamatsu S8664-55) with $5\times5$~mm$^2$ active area and $75\%$ quantum efficiency were glued on the rear face of the crystals using MeltMount 1.7 thermal plastic adhesive. The low gain of APDs ($\sim 200$) was compensated with custom made preamplifier boards, which provide factor 2000 amplification of the ADP signal.

\begin{figure}[]
\begin{center}
{\small
\includegraphics[width=7cm]{figures/ecal-module-schematic.png}
\caption{Schematic vied of an ECal module.}
\label{fig:ecal-module}
}
\end{center}
\end{figure}

@@

During the test run ECal was mounted downstream of the analyzing dipole magnet at the distance of about $147$~cm from the 
upstream edge of the magnet. 
%The two ECal modules are positioned just above and below the ECal vacuum chamber as shown in bottom panel of Fig.~\ref{fig:ecal}. 
%The primary electron beam, scattered electrons and bremsstrahlung photons pass through the system in vacuum. 
The gap between innermost edge of the crystals of two parts was $7.4$~cm, centered on the beam plane. The temperature stability of the system during the test run was better than $1^\circ~$F, the temperature uniformity inside the box was better than  $4^\circ~$F.

@@

\subsection{Signal readout}
\label{sec:fadc}

ECal signal readout was organized as follows: the amplified APD signal from the motherboard is sent to a $2:1$ signal splitter. After the split, 2/3 of the signal is sent a discriminator then to a Time-to-Digital 
Converter (TDC). The 1/3 of the signal is fed to a single channel of 16-channel JLAB Flash Amplitude-to-Digital Converter (FADC250), see Fig.\ref{fig:fadc} \cite{fadc}.  
\begin{figure}[t]
\begin{center}
{\small
\includegraphics[width=7cm]{figures/FADC250_Photo_001.jpg}
\caption{A Jefferson Lab FADC250 VXS module.}
\label{fig:fadc}
}
\end{center}
\end{figure}
Two 20-slot VXS crates with 14 (for top ECal) and 13 (for bottom ECal) FADC bards were employed for the test run.

The FADCs store 12-bit digitized samples at 250~MHz in 8~$\mu$s deep pipelines. 
When a trigger is received, the appropriate part of the pipeline is accessed for the readout. HPS used FADCs in the integration mode. A predefined number of samples of the signal that exceeds a predefined threshold have been summed and stored in 17-bit register for readout. The number of samples for integration was defined by two parameters that could be set separately, the number of samples integrated before the signal crossed threshold ($NSB$) and  the number of samples integrated after the signal crossed threshold ($NSA$), see Fig.\ref{fig:hps_trigger_data}. 
This scheme significantly compresses the data input 
to the FADC. During data analysis, a pedestal value is subtracted to obtain the actual summed energy.

For various reasons only 385 out of 442 modules (87\%) were operational during the test run. The readout threshold was set to $70$ MeV, the number of samples before and after the threshold crossing were NSB=5 and NSA=30, respectively.

\begin{figure}[]
\begin{center}
{\small
\includegraphics[width=8cm]{figures/fadc_datapath_diagram.png}
\caption{\small{FADC data paths}}
\label{fig:hps_trigger_data}
}
\end{center}
\end{figure}

@@

The HPS test run trigger system is designed to efficiently select \ee{} events by 
using information from the ECal . The trigger looks for time coincidences of clusters in the top and bottom half 
of the ECal.  The trigger system can be broken down into three sections (see Fig.~\ref{fig:hps_trigger_cal}):
 \begin{itemize}
 \item FADC (pulse finding): samples and digitizes the signal pulses from each detector channel. Sends the measured pulse energy and arrival time to the Crate Trigger Processor (CTP),
\item CTP (cluster finding): groups FADC pulses from each half of the ECal into clusters. The cluster energy and arrival time are sent to the Sub-System Processor (SSP),
 \item SSP (cluster pair finding): searches for time coincidence of pairs of clusters from the top and bottom half of the ECal and applies topological selections.
 \end{itemize}
 \begin{figure}[t]
\begin{center}
{\small
 \includegraphics[width=8cm]{figures/hps_trigger_cal}
\caption{Block diagram of the ECAL trigger system consisting of the FADC that samples and digitizes signals for each detector channel and sends them for cluster finding in the CTP. The CTP clusters are sent to the SSP where the final trigger decision is taken based on pairs of clusters in both halves of the ECal. The decision is sent to the Trigger Supervisor (TS) that generates the necessary signals to readout the sub-detectors.}
 \label{fig:hps_trigger_cal}
}
\end{center}
 \end{figure}
The time coincidence window of pairs of clusters in the top and bottom half of the ECal are programmable 
with 4~ns resolution.

As described above, the first stage of the trigger logic is incorporated into the 
FPGA's on the FADC boards. The trigger  
integrates using a "time-over-threshold" window, see Fig. \ref{fig:hps_trigger_data}.
For the trigger path, every channel has: 
\begin{itemize}
% \item number of samples integrated before the threshold crossing (NSB),
 %\item number of samples integrated after the  threshold crossing (NSA),
 \item trigger threshold, measured in ADC counts, 
 \item an offset, 
 \item a conversion factor that converts the ADC counts into 5-bits energy in MeV, from $50$ to $\sim1500$~MeV, 
 \item an energy discriminator threshold (minimum energy cutoff).
 \end{itemize}
Note that the threshold for the trigger path can be set independently from the readout threshold.
The offset is subtracted from ADC samples before they are integrated.  
The values reported to the CTP are the 5-bit pulse energy and 3-bit time information at which the 
pulse crossed the threshold. Data for every channel is sent to CTP every 32~ns. With the available 
3-bit time information, the CTP looks for time coincidence of crystal signals with 8~ns resolution (limit is 4~ns).
The cluster finding algorithm is fast and makes use of the parallel processing nature of FPGA's by 
simultaneously searching for 125 clusters, up to 3x3 in size, across the calorimeter crystal array and performs 
the following tasks:
\begin{itemize}
\item Adds energy from hits together for every 3x3 square of channels in ECal.
\item Hits are added together if they occur (leading edge) within a programmable number of 4~ns clock cycles of each other (HPS test run used 8~ns coincidence time interval).
\item If the 3x3 energy sum is larger than the programmable cluster energy threshold and the sum is greater than any neighboring 3x3 windows, the CTP reports the cluster parameters to the SSP. 
\end{itemize}
The CTP evaluates all hits in its half of the calorimeter every 4~ns. A programmable time window is used to 
allow hits that are out of time with each other to be considered as part of a cluster sum. This is done by 
reporting hits when they occur and then reporting them again for the next $N$ number of 4~ns clock cycles, 
where $N \in [0,7]$. This is useful to deal with skew and jitter that develop from the detector, cabling, and 
electronics. As described above, the CTP only selects the 3x3 window with the highest energy sum of its 
neighbors. This filtering is applied to deal with overlapping clusters and cases where the cluster is larger than 
a 3x3 window.

@@

During the test run, for the most part trigger system run in single cluster mode and was tested to work for up to 20 kHz trigger rates.

\subsection{Data Acquisition and Online Computing}
\label{sec:daq}
For the ECal, every VXS crate contains a Readout Controller (ROC)
that collects digitized information, processes it, and sends it on to the Event Builder (EB). The ROC is a single 
blade Intel-based CPU module running DAQ software under CentOS Linux OS. For the SVT ATCA system, 
the ROC application runs on an embedded processor situated on the ATCA main board. The EB assembles 
information from the SVT and ECal ROCs into a single event which is passed to the Event 
Recorder (ER) that writes it to a RAID5-based data storage system capable of handling up to 100 MB/s. The 
EB and other critical components run on multicore Intel-based multi-CPU servers. The DAQ network system is 
a Foundry router providing high-speed connections between the DAQ components and to the JLab 
computing facility. The SVT ROC, which must handle large data volumes, has a 10 Gbit/s link to the Foundry 
router, while a 1 Gbit/s link is adequate for the ECal. A 10 Gbit/s uplink to the JLab 
computing facility is used for long-term storage.

@@

%step Of 442 modules, 39 were disabled or disconnected and were not read out by the DAQ. 13 of these were not read out because of a shortage of FADC readout boards (see Sec.~\ref{sec:fadc}).The remainder either had no HV bias on the APD, or were disabled in the FADC software due to noise. In the data, we identified two types of abnormal channels. One FADC was not sending trigger signals correctly, resulting in low efficiency. This affected the 13 channels read out by that FADC. 5 channels were diagnosed as noisy because they  had a high incidence of hits out of coincidence with the trigger. A large number of channels were originally misidentified as noisy because they had much higher hit occupancy than neighboring channels. Gain calibration shows that these channels have high gain (and thus lower energy threshold) but are otherwise normal. The abnormal channels were ignored in analysis in order to simplify comparison with Monte Carlo. This leaves 385 useful channels---87\% of the ECal.

The gain of the individual ECal channels has been calibrated using the SVT measurement of track momentum and comparison to Monte Carlo simulation. The integrated pulse of the each channel was converted to energy by first subtracting the pedestal then applying a gain or conversion factor to convert ADC counts to energy in MeV. The pedestals have been measured using special runs where FADCs sampled APD signals every 4~ns and for each trigger a time window of 100 samples have been recorded. Pedestal levels were determined by looking at a part of the 
window before the actual hit. Modules above a threshold are clustered using a simple algorithm described in Sec.~\ref{sec:trigger}. The ECal was operated 
with an effective readout threshold of 73~MeV. The noise level of the ECal modules were measured to be approximately 15~MeV.

To compare data and simulation, all inoperable or noisy channels in SVT and ECal have been disabled in the simulation, so any efficiency or bias effects that affect the real data should be reflected 
in the simulation as well. Using a formula to compute the ``weighted E/p'' for a crystal, representing the average E/p for clusters that include the crystal
% $\frac{\sum_j w_{j,i}}{\sum_j\frac{P_j}{E_j}w_{j,i}}$
, and iteratively adjust the gains until the weighted E/p is equal for test run data and simulation.
\begin{figure}[]
\begin{center}
{\small
	\includegraphics[width=7cm]{figures/eoverp_simulation}
	\caption{Weighted E/p from Monte Carlo simulation after gain calibration.}
	\label{fig:gains}
}
\end{center}
\end{figure}

These gains can then be used to convert from ADC counts in a channel to the energy deposited into that ECal 
crystal.
The other information needed to find the energy of an incident particle is the sampling fraction---the ratio of 
energy read out from cluster to energy of an incident particle. The conventional sampling fraction, 
the fraction of incident energy that is deposited in the cluster, is approximately 0.9  for the ECal (somewhat less at the 
edges). For our readout, there is additional energy lost because crystals under the threshold are not 
read out. The weighted E/p used in calibration (see Figure \ref{fig:gains}) is an approximate measurement of sampling fraction, but the sampling fraction is energy-dependent because of the effect of readout threshold. 
A full computation of sampling fraction can be done using simulation.

@@

To eliminate trigger bias in checking the trigger decision, we use a tag and probe method. To check trigger 
performance in one half of the ECal, we tag events where there was a trigger in the other half, and exactly one 
probe cluster in the ECal half under test. 
We then measure trigger efficiency (proportion of tagged events where there was a trigger) as a function of 
ADC counts and energy of the probe cluster.
These turn-on curves are shown for the top half of the ECal in Fig.~\ref{fig:turnon}.

\begin{figure}[ht]
\begin{center}
{\small
	\includegraphics[width=7cm]{figures/top_turn_on__bottom_tag}
	\includegraphics[width=7cm]{figures/top_turn_on__bottom_tag_E}
	\caption{Trigger turn-on as a function of probe cluster ADC counts (top) and probe cluster energy in MeV
	(bottom). Both plots are for the top half of the ECal; bottom is similar. Energy is not corrected for sampling 
	fraction.}
\label{fig:turnon}
}
\end{center}
\end{figure}
The trigger threshold is seen to be 1280 ADC counts as expected. The threshold is not perfectly sharp in this 
analysis because of uncertainties in the conversion from readout to trigger hits but based on comparisons 
with Monte Carlo simulation we believe the trigger worked exactly as specified. 
The trigger threshold in terms of cluster energy is very uneven for two reasons; gain variations between 
different ECal crystals lead to threshold variations and the nonlinearity of the time-over-threshold integral 
means that the effective threshold is higher for clusters that span multiple crystals.
Overall the trigger appears to have functioned exactly as intended. Changes planned for the next run 
(constant integration window and per-crystal gain calibration constants for the trigger) will solve both of the 
issues that led to threshold variations in the test run.
